Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-07 paper/disc20clustering.tex
the intuition paragragh and more

 \newcommand{\OPT}{\textsf{OPT}\xspace}
 \newcommand{\ALG}{\textsf{ALG}\xspace}
 \newcommand{\PPL}{\textsf{PPL}\xspace}
 \newcommand{\OBRP}{BRP}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\textsf{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg

@@

We charge both \OPT and \RM whenever \RM collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to \OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge \OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
%This observation suggests that we can't charge \OPT more than $\alpha$ per collocation.
%E.g.,
%OPT can collocate two costly pairs with a single swap if
%it has both pairs on the cut of  the same two clusters.
%Later, \RM collocates those pairs with two swaps.
Hence,
whenever \OPT collocates a pair,
we charge it only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on Alg).

Observe that for requests in
$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $ 
and the entire duration of the input sequence (i.e., $(0,m]$),
OPT must pay a total cost
%We denote the cost w.r.t~$r_t$ by
$\mathit{OPT} (R_t) \geq 
	c_{R_t} := \min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$,
for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
regardless of its choices.
Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
%
However,
we can charge this cost to \OPT only at $t=t_0$,
i.e.~the first time \RM collocates $\{u,v\}$,
and not at any consequent time that \RM collocates $\{u,v\}$.
That is,
we charge \OPT a cost incurred by requests in
$\sigma_{\{u,w\}}$ or $\sigma_{\{u,v\}}$,
if and only if  \OPT incurs the cost during  intervals $(t'', t']$ or $(t', t]$,
respectively.	
Latter,
this ensures us that we do not charge \OPT for a cost more than twice.
%%
We use notations defined so far in the proof that follows.

@@

	If $t$ is not the first time that \RM collocates $\{u,v\}$ then
	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
	in order to collocate some (intersecting) pair
	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
	e.g.,
	$\{x,y\}=\{u,w\}$.
	Else,
	$t$ is the first time that \RM collocates $\{u,v\}$ and let $t' := 0$.
	Similarly,
	if $t' > 0$ is not the first time that \RM  collocates $\{u,w\}$ 
	then let $0 < t'' < t'$ be the latest time before $t'$ when \RM separates $\set{u,w}$.
	Else,
	$t'$ is the first time that \RM collocates $\{u,w\}$ and we let $t''=0$.
	%

	First,
	we bound the costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by Alg,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound the cost incurred by Alg
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for these requests $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is

	\begin{align*}	%\label{eq:costAlg}
	\mathit{ALG} (\sigma)
	=
	\sum_{ t \in T}(\lambda + 2\alpha) +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
	\leq
	\sum_{ t \in T}(\lambda + 2\alpha) +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
	=
	\sum_{ t \in T} 3\alpha +
	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
	.
	\end{align*}

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with cost 0.
	Therefore in this case,
	$\mathit{OPT} (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$ in an analogous way.
	Then,
	$\mathit{OPT} (R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \mathit{OPT} (\sigma_{\{u,v\}}) $.
	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
	Latter,
	 we argue that \OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		let $\OPT_1 (R_t) := 2\alpha$.
		\item
		\OPT has one of the pairs  already collocated and the other pair separated,
			  prior to and during their respective intervals.
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$
		and let $\OPT_2 (R_t) := \lambda$.
		\item
		\OPT has one of the pairs already collocated prior to its interval,
		and it collocates the other pair during the respective interval.
		Then it pays 0 for one pair and $\alpha$ for the other and 
		let $\OPT_3 (R_t) := \alpha$.
	\end{enumerate}
%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
	If \OPT behaves in one of these cases,
	then
	$\mathit{OPT} (R_t) \geq
		\min{\{ \OPT_1(R_t), \OPT_2(R_t), \OPT_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	\OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
	Therefore in any case,
	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.

@@

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\mathit{Cost}_{OPT} ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
	and  
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by Alg,
	and for final costs at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in [1,m] } \mathit{OPT} (R_t) +
	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

	\begin{align*} 	%\label{eq:costOPT}
		\mathit{OPT} (\sigma)
		&=
		\frac{1}{2}
		\Big(
		\sum_{ t \in T} \mathit{OPT} (R_t) 
		+ \sum_{\{u,v\}}\mathit{OPT} (R^*_{\{u,v\}})
		\Big)	\\
		&\leq
		\Big(
		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
		\Big)		
		=
		\frac{1}{2}		
		\Big(
		\sum_{ t \in T} \alpha  
		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
		\Big).
	\end{align*}

>>>>

 \newcommand{\OPT}{\textsf{OPT}\xspace}
  \newcommand{\OPTM}{\mathit{OPT}}
 \newcommand{\ALG}{\textsf{ALG}\xspace}
 \newcommand{\PPL}{\textsf{PPL}\xspace}
 \newcommand{\OBRP}{BRP}
 \newcommand{\PPOBRP}{PP-BRP}
 \newcommand{\dist}{\textsf{dist}}
 \newcommand{\TAlg}{{\ensuremath{\textsf{ALG}_{3}}}\xspace}
  \newcommand{\RM}{\textsf{RM}\xspace} % rematching alg

@@

We charge both \OPT and \RM whenever \RM collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to \OPT for the first time.
\RM collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$,
while OPT may save some costs by collocating multiple pairs at once, 
 before they inflict too much communication cost.
 Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
Hence,
whenever \OPT collocates a pair,
we charge it only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).

\textbf{The charging Scheme.}
Consider two  pairs that share a same node, 
i.e.~\emph{intersecting pairs},
and the set of requests that causes the collocation of both pairs,
at some  times during  $[1,m]$.
%i.e.~the entire duration of the sequence.
Observe that \OPT must pay a non-zero cost
for these requests over the entire $\sigma$,
since it cannot have both pairs initially collocated.
%More precisely,
%OPT  pays a total cost at least
%	$\min{ \{ 2\lambda, \lambda+\alpha, 2\alpha \} }$ for, respectively,
%	serving requests to both pairs remotely,
%	serving requests to one pair remotely and collocating the other,
%	or collocating both pairs. 
%
%for serving (some of) these requests remotely or (and) collocating (one of) the pairs,
%regardless of its choices.
%Then the cost per pair, in particular the cost w.r.t.~$\{u,v\}$, is
%$\mathit{OPT} (\sigma_{\{u,v\}}) \geq c_{R_t} / 2$.
%
However,
we can charge the whole cost to \OPT only the first time \RM collocates a pair,
and not at any consequent time when \RM collocates it a second time.
Otherwise,
 \OPT is possibly charged for a same cost repeatedly.
For this reason,
we charge \OPT a cost inflicted by a pair,
if and only if  \OPT incurs the cost some time since the last time \RM separated the pair.
%%

@@

	If $t$ is not the first time that \RM collocates $\{u,v\}$ then
	let $0 < t' < t$ be the latest time when \RM separates $\set{u,v}$
	in order to collocate some intersecting pair
	$\{x,y\} \neq \{u,v\}, \{x,y\} \cap \{u,v\} \neq \emptyset$, 
	e.g.,
	$\{x,y\}=\{u,w\}$.
	Else,
	$t$ is the first time that \RM collocates $\{u,v\}$ and let $t' := 0$.
	Similarly,
	if $t' > 0$ is not the first time that \RM  collocates $\{u,w\}$ 
	then let $0 < t'' < t'$ be the latest time before $t'$ when \RM separates $\set{u,w}$.
	Else,
	$t'$ is the first time that \RM collocates $\{u,w\}$ and we let $t''=0$.
	%

	First,
	we bound  costs incurred by \RM for requests that
	lead to the collocation of $\{u,v\}$ at time $t \in T$, where
	$T := \{ i \in [1,m] ~\vert~ \exists \{x,y\}: C^{i}_{\{x,y\}} = \lambda \}$
	is the set of times when \RM performs a collocation.
	By definitions of $t$ and $t'$,
	the overall cost of requests in $\sigma_{\set{u,v}}$ incurred by \RM,
	i.e., the total cost of remote serving
	and the moving cost, is
	$\lambda + 2\alpha$.	
	Next,
	we bound costs incurred by \RM
	for requests that do not lead to collocations until the  end of the sequence at $t=m$.
	Assume $\{u,v\}$ is not collocated at $t=m$
	and $0 < C^{m}_{ \{u,v\} } < \lambda $,
	which means \RM pays $C^{m}_{ \{u,v\} }$
	for  requests in $\sigma_{\set{u,v}}(t',m]$.
    Then the overall cost to \RM is
$	\mathit{RM} (\sigma)
=
\sum_{ t \in T}(\lambda + 2\alpha) +
\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
$.
%	\begin{align*}	%\label{eq:costAlg}
%	\mathit{RM} (\sigma)
%	=
%	\sum_{ t \in T}(\lambda + 2\alpha) +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
%	\leq
%	\sum_{ t \in T}(\lambda + 2\alpha) +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}	
%	=
%	\sum_{ t \in T} 3\alpha +
%	\sum_{\{u,v\}} C^{m}_{\{u,v\}}
%	.
%	\end{align*}

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with cost 0.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$ in an analogous way.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
	Latter,
	 we argue that \OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		let $\OPTM_1 (R_t) := 2\alpha$.
		\item
		\OPT has one of the pairs  already collocated and the other pair separated,
			  prior to and during their respective intervals.
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$
		and let $\OPTM_2 (R_t) := \lambda$.
		\item
		\OPT has one of the pairs already collocated prior to its interval,
		and it collocates the other pair during the respective interval.
		Then it pays 0 for one pair and $\alpha$ for the other and 
		let $\OPTM_3 (R_t) := \alpha$.
	\end{enumerate}
%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
	If \OPT behaves in one of these cases,
	then
	$\mathit{OPT} (R_t) \geq
		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	\OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
	Therefore in any case,
	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.

@@

	Next, we sum up all costs incurred by \OPT.
	By definitions of $R_t$ and $R^*_{\{u,v\}}$, we have either
	$R_{t'} \cap R_t = \sigma_{\{u,w\}}$ or
	$R_{t'} \cap R^*_{\{u,v\}} = \sigma_{\{u,w\}}$
	(\mahmoud{See figure \ref{fig:doubleCount}}). 
	This means,
	$\OPTM ( \sigma_{\{u,w\}})$
	is counted at most twice in each of  the expressions
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R_t)$
	and  
	$\mathit{OPT} (R_{t'}) + \mathit{OPT} (R^*_{\{u,v\}})$.
	Hence,
	for all collocations performed by \RM,
	and for final costs at $t=m$,
	\OPT pays at least 
	$\frac{1}{2}(
	\sum_{ t \in T } \mathit{OPT} (R_t) +
	\sum_{\{u,v\}} \mathit{OPT} (R^*_{\{u,v\}})
	) $.
	Then,
	the total cost to \OPT is

	\begin{align*} 	%\label{eq:costOPT}
		\mathit{OPT} (\sigma)
		&=
		\frac{1}{2}
		\Big(
		\sum_{ t \in T} \mathit{OPT} (R_t) 
		+ \sum_{\{u,v\}}\mathit{OPT} (R^*_{\{u,v\}})
		\Big)	\\
		&\geq
		\Big(
		\sum_{ t \in T} \min{ \{ \lambda, \alpha \}}  +
		\sum_{\{u,v\}} \min{ \{C^{m}_{\{u,v\}} , \alpha \} } 
		\Big)		
		=
		\frac{1}{2}		
		\Big(
		\sum_{ t \in T} \alpha  
		+ \sum_{\{u,v\}} C^{m}_{\{u,v\}}
		\Big).
	\end{align*}
